Improving the Image Acquisition Rate \\ of an Atomic Force Microscope \\ Through Sub-sampling and Reconstruction

### 2019-07-01 latex/afm-cs-main.tex
Final commit at submission.

Our previous work on the $\mu$-path pattern, using theoretical
calculations, simulations, and a preliminary implementation,
demonstrates that the approach can achieve scanning time
reduction while maintaining faithful image reconstruction
\cite{maxwell2014compressed,Luo:2015tu, braker_hardware_2018}. 
The main contributions of this paper relative to that past work are as follows.
% \begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
\begin{itemize}
\item We introduce a new reconstruction algorithm which is designed to
minimize the artifacts arising from the structure of the $\mu$-path
scanning pattern. This algorithm is described and compared to existing
reconstruction methods through simulations in Sec. \ref{sec:reconstructionMethods}.
\item In our prior experimental prototype \cite{braker_hardware_2018}, we used very basic integral controllers for all three axes for both raster and $\mu$-path scans. Here, we design compensators for each axis using standard loop-shaping techniques (Sec.~\ref{sec:control}), which result in a substantial increase in closed-loop bandwidth (e.g., 50 Hz to 450~Hz for the $Z$-axis).
We show in Sec.~\ref{sec:ff_control} that one of the limitations of $\mu$-path scanning is the discontinuous $X$-axis reference (e.g., in moving between $\mu$-paths) which excites the cross-coupling modes between the $X$ and $Z$ axes, an effect which neither our prior experimental work nor simulation studies accounted for. We mitigate the cross coupling with a feed-forward design.
\item We experimentally compare (Sec. \ref{sec:results}) $\mu$-path scanning to not only full density raster scans but also to coarse raster scans which are interpolated to a full image. This is in contrast to typical sub-line sampling which takes scan lines at random. The clear advantage to a coarse raster scan is that its implementation only requires adjustments to the data collection and post-processing. 
  % it requires no modification to the instrument, only to the post-processing.
\item The bulk of this work is based on experiment, not simulation, in contrast to prior work by ourselves \cite{luo2015comparison,maxwell2014compressed} and others \cite{han_optimal_2018,jensen_reconstruction_2013,oxvig_structure_2017}. Our main focus here is on delineating the practical issues which must be addressed to effectively realize $\mu$-path scanning, which has only seen limited attention in the past.
% \end{enumerate}
\end{itemize}

@@

\begin{figure}[t!]
  \centering \includesvg[width=1\columnwidth]{figures/cs_cycle.svg}
  \caption{Several CS cycles. Each state is indicated by color.}
  \label{fig:cs_cycle}
\end{figure}

@@

In general, the $X$-axis positioning requirements are more stringent
for both CS and raster scanning. Due to the step inputs the $X$-axis
must track, we have found that compensating hysteresis and creep to be
particular helpful during $\mu$-path scanning. Thus, the $X$-axis
compensator also includes an inverse creep and hysteresis model. The
hysteresis model is a modified Prandtl-Ishlinskii type model and is
based on the work of \cite{kuhnen_modeling_2003}, while the drift
compensation models the piezo creep as a simple second-order transfer
function with real poles and zeros whose inverse is part of $D_X$.
In Fig.~\ref{fig:afm_bd_dinv}, the hysteresis inversion is indicated
by $\mathcal{H}_X^{-1}[\cdot]$. We
have found that the inclusion of these extra compensators limits the
overshoot of the $X$-axis during larger moves between different
$\mu$-paths (e.g., without them, the $X$-axis trajectory in
Fig.~\ref{fig:with_ff} would show noticeable overshoot). More details
on fitting the creep and hysteresis models can be found in our other
work \cite{braker_afmmpc_2019}.

@@

\subsection{Comparison to Sub-line Sampling} \label{sec:cp_subline}
Many sub-line sampling papers assume that the lines to sample are
selected randomly \cite{han_optimal_2018, Luo_nano_2015}. An even
simpler alternative to $\mu$-path sampling is simply to take a
standard raster scan with fewer lines. This is the technique we
consider in this subsection, by scanning images with 128 and 64 lines.
For each of these scenarios, we first divide the data for each line
into 512 bins. This gives an image that is 128 (resp., 64) pixels by
512 pixels, corresponding to 25\% (resp., 12.5\% sampling). We then linearly interpolate along each column to produce
a 512 by 512 pixel image. Again, we take these coarser raster scans at
1.0, 2.5, 5.0, 8.0 and 10.0 Hz. A subset of the resulting images are shown in
Fig.~\ref{fig:results_subline_images}.

@@

% Note the defect on the hole located in the second column, sixth row
% from the top (of the grating holes). The CS-scans do a poor job
% detecting this, while both coarse raster scans do so. However, the
% other artifacts in the 64 line scan obscure the defect. It is important to note, however, that the appearance (or lack thereof) of such small features in the image is in large part a random occurrence. In our scans, the line scanner happened to pass through this region and capture the appropriate data while the $\mu$-path pattern used did not. 
% Except that it passed through this region for all of the sub-line scans, which did not scan exactly the same area...

Fig.~\ref{fig:improve_512} compares the amount of speed improvement of
the different sub-sampling schemes over a full raster scan with 512
lines as a function of scan rate. As far as pure speed goes, the CS
scan at 12.5\% is the fastest for a rate of 1~Hz, though the 64 line
raster scan is close. At the same time, the image quality of the 1~Hz,
12.5\% CS scan is better than the 1~Hz, 64 line raster scan.
This speed advantage quickly evaporates for faster scan rates.
Indeed, for rates beyond about 3~Hz, the 64 line scan is faster than the 12.5\% CS scan and
the 128 line scan is faster than the 25\% CS scan. 
In general, the artifacts of the CS reconstructions are different than the interpolated raster scans.
CS appears to preserve the edges of the holes better than sub-line sampling,
while sub-line sampling has less noise in the flat areas of the specimen.

>>>>

Our previous work on the $\mu$-path pattern, using theoretical
calculations, simulations, and a preliminary implementation,
demonstrates that the approach can achieve scanning time
reduction while maintaining faithful image reconstruction
\cite{maxwell2014compressed,Luo:2015tu, braker_hardware_2018}. 
The main contributions of this paper relative to that past work are as follows.
% \begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
\begin{itemize}
\item We introduce a new reconstruction algorithm which is designed to
minimize the artifacts arising from the structure of the $\mu$-path
scanning pattern. This algorithm is described and compared to existing
reconstruction methods through simulations in Sec. \ref{sec:reconstructionMethods}.
\item In our prior experimental prototype \cite{braker_hardware_2018}, we used very basic integral controllers for all three axes for both raster and $\mu$-path scans. Here, we design compensators for each axis using standard loop-shaping techniques (Sec.~\ref{sec:control}), which result in a substantial increase in closed-loop bandwidth (e.g., 50 Hz to 450~Hz for the $Z$-axis).
We show in Sec.~\ref{sec:ff_control} that one of the limitations of $\mu$-path scanning is the discontinuous $X$-axis reference (e.g., in moving between $\mu$-paths) which excites the cross-coupling modes between the $X$ and $Z$ axes, an effect which neither our prior experimental work nor simulation studies accounted for. We mitigate the cross coupling with a feed-forward design.
\item We experimentally compare (Sec. \ref{sec:results}) $\mu$-path scanning to not only full density raster scans but also to coarse raster scans which are interpolated to a full image. This is in contrast to typical subline sampling which takes scan lines at random. The clear advantage to a coarse raster scan is that its implementation only requires adjustments to the data collection and post-processing. 
  % it requires no modification to the instrument, only to the post-processing.
\item The bulk of this work is based on experiment, not simulation, in contrast to prior work by ourselves \cite{luo2015comparison,maxwell2014compressed} and others \cite{han_optimal_2018,jensen_reconstruction_2013,oxvig_structure_2017}. Our main focus here is on delineating the practical issues which must be addressed to effectively realize $\mu$-path scanning, which has only seen limited attention in the past.
% \end{enumerate}
\end{itemize}

@@

\begin{figure}[t!]
  \centering \includesvg[width=1\columnwidth]{figures/cs_cycle.svg}
  \caption{Several CS cycles. Each state is indicated by color. The deflection is shifted so its free value is zero.}
  \label{fig:cs_cycle}
\end{figure}

@@

In general, the $X$-axis positioning requirements are more stringent
for both CS and raster scanning. Due to the step inputs the $X$-axis
must track, we have found that compensating hysteresis and creep to be
particular helpful during $\mu$-path scanning. Thus, the $X$-axis
compensator also includes an inverse creep and hysteresis model. The
hysteresis model is a modified Prandtl-Ishlinskii type model and is
based on the work of \cite{kuhnen_modeling_2003}. The drift
compensation models the piezo creep as a simple second-order transfer
function with real poles and zeros whose inverse is part of $D_X$.
In Fig.~\ref{fig:afm_bd_dinv}, the hysteresis inversion is indicated
by $\mathcal{H}_X^{-1}[\cdot]$. We
have found that the inclusion of these extra compensators limits the
overshoot of the $X$-axis during larger moves between different
$\mu$-paths (e.g., without them, the $X$-axis trajectory in
Fig.~\ref{fig:with_ff} would show noticeable overshoot). More details
on fitting the creep and hysteresis models can be found in our other
work \cite{braker_afmmpc_2019}.

@@

\subsection{Comparison to subline Sampling} \label{sec:cp_subline}
Many subline sampling papers assume that the lines to sample are
selected randomly \cite{han_optimal_2018, Luo_nano_2015}. An even
simpler alternative to $\mu$-path sampling is simply to take a
standard raster scan with fewer lines. This is the technique we
consider in this subsection, by scanning images with 128 and 64 lines.
For each of these scenarios, we first divide the data for each line
into 512 bins. This gives an image that is 128 (resp., 64) pixels by
512 pixels, corresponding to 25\% (resp., 12.5\% sampling). We then linearly interpolate along each column to produce
a 512 by 512 pixel image. Again, we take these coarser raster scans at
1.0, 2.5, 5.0, 8.0 and 10.0 Hz. A subset of the resulting images are shown in
Fig.~\ref{fig:results_subline_images}.

@@

% Note the defect on the hole located in the second column, sixth row
% from the top (of the grating holes). The CS-scans do a poor job
% detecting this, while both coarse raster scans do so. However, the
% other artifacts in the 64 line scan obscure the defect. It is important to note, however, that the appearance (or lack thereof) of such small features in the image is in large part a random occurrence. In our scans, the line scanner happened to pass through this region and capture the appropriate data while the $\mu$-path pattern used did not. 
% Except that it passed through this region for all of the subline scans, which did not scan exactly the same area...

Fig.~\ref{fig:improve_512} compares the amount of speed improvement of
the different sub-sampling schemes over a full raster scan with 512
lines as a function of scan rate. As far as pure speed goes, the CS
scan at 12.5\% is the fastest for a rate of 1~Hz, though the 64 line
raster scan is close. At the same time, the image quality of the 1~Hz,
12.5\% CS scan is better than the 1~Hz, 64 line raster scan.
This speed advantage quickly evaporates for faster scan rates.
Indeed, for rates beyond about 3~Hz, the 64 line scan is faster than the 12.5\% CS scan and
the 128 line scan is faster than the 25\% CS scan. 
In general, the artifacts of the CS reconstructions are different than the interpolated raster scans.
CS appears to preserve the edges of the holes better than subline sampling,
while subline sampling has less noise in the flat areas of the specimen.
